A novel route to the spontaneous formation of porous crystals via viscoelastic phase separation

### 2016-03-21 crystal-gel/nature/Supplementary.tex
Small additions.

\begin{figure}
\includegraphics{local_phi12.pdf}
\caption{Late stage of the evolution of local volume fraction around colloidal particles having 12 nearest neighbours. Samples are shown with the same color code as in Fig.1 of the main text. In particular red circles correspond to the crystalizing sample at $\phi=0.33$ and $cp=0.38\,$mg/g.}
\label{fig:local_phi12}
\end{figure}

@@

Our two constraints is (i) to have the final local volume fraction of the 12-neighboured particles in the crystallising sample to be equal to the theoretical equilibrium crystal volume fraction and (ii) to have the theoretical spinodal line to correspond to the gel region boundary. Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=80$~nm, $\sigma = 2.25\,\mu$m and a crystal at $0.723$.

@@

@@

\clearpage
\section*{Supplementary Movie 1}
This movie shows the early stage of gelation process of $\phi\approx 0.30$ and $c_p=0.38$ mg/g, from the original confocal images at the same z-position as Fig.~3a. Total duration is 1160 s (530~$\tau$). We quickly started scanning the sample after exchanging the salt-reservoir with the solvent including salt. At the beginning, salt had not reached the scanning volume and colloids were dispersed by electrostatic repulsion. At 0:02 in the movie, salt diffused into the scanning volume and screened the surface charges of colloids. Then, colloids immediately initiated phase separation by short-ranged attractions.

>>>>

\begin{figure}
\includegraphics{local_phi12.pdf}%
\caption{Late stage of the evolution of local volume fraction around colloidal particles having 12 nearest neighbours. Samples are shown with the same color code as in Fig.1 of the main text. In particular red circles correspond to the crystalizing sample at $\phi=0.33$ and $cp=0.38\,$mg/g.}%
\label{fig:local_phi12}%
\end{figure}

@@

Our two constraints are (i) to have the final local volume fraction of the 12-neighboured particles in the crystallising sample to be equal to the theoretical equilibrium crystal volume fraction and (ii) to have the theoretical spinodal line to correspond to the gel region boundary. Starting from our initial estimates of $\sigma$ and $R$ we iteratively converge to $R=80$~nm, $\sigma = 2.25\,\mu$m and a crystal at $0.723$.

@@

\section*{Capillary experiment}

\begin{figure*}[!t]
\includegraphics{capillary.pdf}
\caption{\textbf{Capillary experiment reproducing the crystal-gel state} Colloid volume fraction is $\approx 0.33$ and polymer concentration is $0.40\,$mg/g. Top row: Confocal slices showing the initial network, crystallising network and final crystal-gel network. Bottom row: Reconstructions from confocal coordinates. Particles are drawn to scale and coloured according to their phase. Red: crystal; white: liquid; green: gas. To highlight the crystalline regions, liquid and gas particles are drawn to a tenth of the crystalline particle size.}
\label{fig:capillary}
\end{figure*}

The use of a reservoir of a salt solution through a semi-permeable membrane, allowed us to follow the entire microscopic dynamics in absence of spurious flows, from the first stages of phase separation to the formation of the crystal-gel state. In order to confirm the reproducibility of the crystal-gel state, we conducted experiments in a more traditional capillary setting, where all components of the suspension are pre-mixed in a sealed cell. A typical example is
shown in Fig.~\ref{fig:capillary}, where a system with colloid volume fraction of $\approx 0.33$ and polymer concentration is $0.40\,$mg/g is considered.
The figure shows the time evolution of the sample at three different times, just after phase separation ($t=340\,\tau_\beta$), and during
crystal-gel formation ($t=2600\,\tau_\beta$ and $t=6650\,\tau_\beta$). The top-row shows images from quasi two-dimensional confocal slices, while the bottom-row shows
the three-dimensional reconstruction after single-particle tracking, where crystalline particles are highlighted in both size and color (red).

\section*{Supplementary Movie 1}
This movie shows the early stage of gelation process of $\phi\approx 0.30$ and $c_p=0.38$ mg/g, from the original confocal images at the same z-position as Fig.~3a. Total duration is 1160 s (530~$\tau$). We quickly started scanning the sample after exchanging the salt-reservoir with the solvent including salt. At the beginning, salt had not reached the scanning volume and colloids were dispersed by electrostatic repulsion. At 0:02 in the movie, salt diffused into the scanning volume and screened the surface charges of colloids. Then, colloids immediately initiated phase separation by short-ranged attractions.

@@

### 2016-03-21 crystal-gel/nature/crystal_gel2.tex
Small additions.

The crystal-gel state was reproduced in a sealed cell with all the components of the suspension pre-mixed. Colloid volume fraction was virtually identical at $\approx 0.33$ and polymer concentration was $0.40\,$mg/g. \textcolor{blue}{See Supplementary Figure XXX.}

>>>>

\textcolor{blue}{
The crystal-gel state is reproduced in a sealed cell with all the components of the suspension pre-mixed. Colloid volume fraction is virtually identical at $\approx 0.33$ and polymer concentration is $0.40\,$mg/g. See Supplementary Figure~\ref{S-fig:capillary}.}
